---
title: "STAT433-HW1-README"
author: "Yi Xiao"
date: "2/2/2021"
output: github_document
---
```{r}
library(readr)
library(MASS)
library(rgenoud)
library(anchors)
library(ggplot2)

```

**Read all bridges data at 2019**
```{r warning=FALSE, include = TRUE}
allRecords2019 <- read.csv("2019AllRecordsDelimitedAllStates.csv") 
```

**Briefly check the database**
```{r}
## head(allRecords2019)
```

From the data description pdf, I find:

Bridge ID: Item 8 - Structure Number

Year Built: Item 27 - Year Built

Fips Code: Item 1 - State Code

Condition Rating: Item 58 through 62 

(Item 58 - Deck,

 Item 59 - Superstructure,
 
 Item 60 - Substructure,
 
 Item 61 - Channel and Channel Protection,
 
 Item 62 - Culverts)
 
 
 

**Select Variables**
```{r}
BridgeID <- allRecords2019$STRUCTURE_NUMBER_008
FipCode <- allRecords2019$STATE_CODE_001
YearBuilt <- allRecords2019$YEAR_BUILT_027
CRDeck <- allRecords2019$DECK_COND_058 #CR stands for Condition Rating
CRSuperstructure <- allRecords2019$SUPERSTRUCTURE_COND_059
CRSubstructure <- allRecords2019$SUBSTRUCTURE_COND_060
CRChannel <- allRecords2019$CHANNEL_COND_061
CRCulverts <- allRecords2019$CULVERT_COND_062
```
 

**Write data into a dataframe** 
```{r}
Bridges2019 <- data.frame(
  "BridgeID" = BridgeID,
  "FipCode" = FipCode,
  "YearBuilt" = YearBuilt,
  "ConRDeck" = CRDeck,
  "ConRSuperstructure" = CRSuperstructure,
  "ConRSubstructure" = CRSubstructure,
  "ConRCulverts" = CRCulverts
)
```


Bridge condition is determined by the lowest rating of National Bridge Inventory (NBI) condition ratings for conRDeck, conRSuperstructure, conRSubstructure, conRCulverts.
So, I will create a new variable "ConR", which is the lowest value of columns conRDeck, conRSuperstructure, conRSubstructure, if the three values are Na, then we will use the ConRCulverts as ConR.

First, I change all 'N' value to 0, so we can filter the lowest value faster
```{r}
Bridges2019 = replace.value(Bridges2019, c("ConRDeck","ConRSuperstructure","ConRSubstructure","ConRCulverts"), from = 'N', to = as.integer(0), verbose = FALSE)
```


**Set the four variables ConRDeck, ConRSuperstructure, ConRSubstructure, ConRCulvert to numeric**
```{r}
Bridges2019$ConRDeck = as.numeric(Bridges2019$ConRDeck)
Bridges2019$ConRSuperstructure = as.numeric(Bridges2019$ConRSuperstructure)
Bridges2019$ConRSubstructure = as.numeric(Bridges2019$ConRSubstructure)
Bridges2019$ConRCulverts = as.numeric(Bridges2019$ConRCulverts)
```

**Find the lowest value among the ConRDeck, ConRSuperstructure, ConRSubstructure, ConRCulvert for every bridge**
```{r}
for (i in 1:nrow(Bridges2019)){
  if (isTRUE(Bridges2019$ConRCulverts[i] == 0)){
    Bridges2019$ConR[i] <- min(Bridges2019[i, c("ConRDeck", "ConRSuperstructure", "ConRSubstructure")])
  }else{
      Bridges2019$ConR[i] = Bridges2019$ConRCulverts[i]
    }
}
```


Write a csv file of all bridges in 2019, each bridge has its own BridgeID, FipCode, YearBuilt, Condition Rating (Deck), Condition Rating (Superstructure), Condition Rating (Substructure), Condition Rating (Culverts), and ConR (the lowest value among the four variable desribed just before)
```{r}
write.csv(Bridges2019, "BridgeRecords2019.csv")
```


Find all bridges in Wisconsin at 2019, and we have the FipCode of Wisconsin is 55
```{r}
WIBridges <- Bridges2019[Bridges2019$FipCode == 55,]
head(WIBridges)
```
```{r}
WIBridges$ConR = as.numeric(WIBridges$ConR)
summary(WIBridges$ConR)
```
From the summary, we could see there is a weird ConR value. The condition rating should be 0 - 10 as indicated in the data description.
(In order to make the following plot clear, I would like to delete the row with ConR of 53.9)
```{r}
WIBridges = WIBridges[-c(4030), ] ## this row had index of 4030
```


```{r warning=FALSE, include = TRUE}
ggplot(WIBridges, aes(ConR)) +
  geom_bar() +
  xlab("Condition Rating") +
  scale_x_binned(limits = c(0, 10), breaks = 0:10)
```

From the data description, we know:
If the lowest rating(ConR) >= 7, the bridge is Good (G).
If the lowest rating(ConR) <= 4, the bridge is Poor (P).
If the 4 < lowest rating(ConR) < 7, the bridge is Fair (F).

Based on the bar plot, we can conclude that most bridges in Wisconsin are Fair, and relatively less bridges are Poor.


Then, I want to know whether the age of bridge (Year of Built) has impacts on the Condition Rating of bridges.
```{r}
summary(WIBridges$YearBuilt)
```

```{r  warning=FALSE, include=TRUE}
ggplot(WIBridges, aes(x = YearBuilt, y = ConR)) +
  geom_jitter() +
  xlim(1930, 2020) +
  ylim(0, 10)
```

From the above plot, we can see the shorter the age of the bridge, the higher the condition rating it would have.
